---
title: "eCog_Ant_mound"
author: "R Pienaar"
date: "`r Sys.Date()`"
output: word_document
---

Libraries

```{r}

library(ggplot2)
library(tidyverse)


```


Insert data:
Input is a combined csv from arc that shows individual disks per treatment and block

```{r}

d16 <- read.csv("Arc_Outputs/disks_2016.csv")
d21 <- read.csv("Arc_Outputs/disks_2021.csv")
d23 <- read.csv("Arc_Outputs/disks_2023.csv")
d16$Year <- 2016
d21$Year <- 2021


d16 <- d16 %>% select(Block, Plot, Shape_Area, Year)
d21 <- d21 %>% select(Block, Plot, Shape_Area, Year)
d23 <- d23 %>% select(Block, Plot, Shape_Area, Year)
disks <- rbind(d16, d21, d23)

```

Create total area per treatment block

```{r}

area <- disks %>% 
  group_by(Block, Plot, Year) %>%
  summarise(Total_Shape_Area = sum(Shape_Area))

complete_combinations <- disks %>%
  distinct(Block, Plot, Year) %>%
  expand(Block, Plot, Year)

# Perform a full join with the original dataframe
area <- complete_combinations %>%
  left_join(area, by = c("Block", "Plot", "Year")) %>%
  mutate(Total_Shape_Area = ifelse(is.na(Total_Shape_Area), 0, Total_Shape_Area))

area <- area %>%
  mutate(Burned = substr(Plot, 1, 1),
         Rodents = substr(Plot, 2, 2))

head(area)

```




Compare total area of ant mounds and disks combined between treatments and years

```{r, warning=FALSE}

hist(area$Total_Shape_Area)

area %>% group_by(Year, Plot) %>%
  summarise(ave = mean(Total_Shape_Area),
            se = sd(Total_Shape_Area)/ sqrt(n()))  |>
  ggplot(aes(Year, ave, color = Plot, fill = Plot))+
   geom_point()+
    geom_ribbon(aes(ymin = ave - se, ymax = ave + se), alpha = 0.1, color = NA)+
    geom_smooth(aes(group = Plot), alpha = 0.1) +
  theme_classic() +
  labs(x = "Year",
       y = expression(paste("Average Total Disk Area per Plot  ", (m^2))))+
  theme(text = element_text(size=12, family = "serif"))
    
area$Year <- as.factor(area$Year)
                    
ggplot(data =  area, aes(Year, Total_Shape_Area, fill = Plot))+
  geom_boxplot()+
  theme_classic()

areadiff <- area |> pivot_wider(names_from = Year, values_from = Total_Shape_Area)
areadiff$tot.change <- areadiff$'2023' - areadiff$'2016'

areaaov <- aov(data = areadiff, tot.change ~ Burned + Rodents + Burned * Rodents)
summary(areaaov)

```

Create density per treatment block

```{r}

density <- disks

density <- density %>% group_by(Block, Plot, Year) %>%
  summarise(mean_count = mean(n()))

density <- complete_combinations %>% left_join(density, by = c("Block", "Plot", "Year")) %>%
  mutate(mean_count = ifelse(is.na(mean_count), 0, mean_count))

head(density)
hist(density$mean_count)

```



```{r, warning=FALSE}

density[c("Burned", "Rodents")] <- str_split_fixed(density$Plot, "", 2)

densityaov <- aov(data = density, mean_count ~ Burned + Rodents)
summary(densityaov)

density %>% group_by(Year, Plot) %>%
  summarise(ave = mean(mean_count),
            se = sd(mean_count)/ sqrt(n()))  |>
  ggplot(aes(Year, ave, color = Plot, fill = Plot))+
   geom_point()+
    geom_ribbon(aes(ymin = ave - se, ymax = ave + se), alpha = 0.1, color = NA)+
    geom_smooth(aes(group = Plot), alpha = 0.1) +
  theme_classic() +
  labs(x = "Year",
       y = "Average Number of Disks per Plot")+
  theme(text = element_text(size=12, family = "serif"))



```




Load manual counts for 2021

```{r}

manual_dens_2021 <- read.csv("Density_manual_2021.csv")
raw <- read.csv("C:/Users/ryanp/Documents/Dissertation/Chapter1_Ants/Ant_analysis_manua;/Ant Manuals.csv")
hist(raw$Disk.Diameter..cm.)

manual_dens_2021 <- manual_dens_2021 %>%
  mutate_all(~replace(., . == "BS","BR"))

manual_dens_2021 <- manual_dens_2021 %>%
  mutate_all(~replace(., . == "US","UR"))

manual_dens_2021 <- manual_dens_2021 |> select(Block, Plot, Count, Area)

head(manual_dens_2021)


manual_dens_2021 <- raw

manual_dens_2021 <- manual_dens_2021 |> group_by(Block, Plot) %>%
  summarise(mean_count = mean(n()))

manual_dens_2021 <- manual_dens_2021[-1,] #remove controls

manual_dens_2021 <- manual_dens_2021 %>%
  mutate_all(~replace(., . == "BS","BR"))

manual_dens_2021 <- manual_dens_2021 %>%
  mutate_all(~replace(., . == "US","UR"))


#### Manual large is to do the same as before but to only compare disks above a certain size threshold
# so same analysis repeated
manual_large <- raw[raw$Disk.Diameter..cm.>= 250,] ### I think there might be a size threshold for detecting mounds

manual_large <- manual_large |> group_by(Block, Plot) %>%
  summarise(mean_count = mean(n()))

manual_large <- manual_large[-1,] # remove controls


manual_dens_2021 <- manual_dens_2021 %>%
  mutate_all(~replace(., . == "BS","BR"))

manual_dens_2021 <- manual_dens_2021 %>%
  mutate_all(~replace(., . == "US","UR"))


head(manual_dens_2021)

```

Compare manual 2021 to drone 2021

```{r}

drone_2021 <- disks[disks$Year== 2021,]
drone_2021 <- drone_2021 %>% group_by(Block, Plot) %>%
  summarise(mean_count = mean(n()),
            Total_Shape_Area = sum(Shape_Area))
drone_2021 <- complete_combinations[complete_combinations$Year==2021,] %>% 
  left_join(drone_2021, by = c("Block", "Plot")) %>%
  mutate(mean_count = ifelse(is.na(mean_count), 0, mean_count))

drone_2021 <- drone_2021 %>% replace(is.na(.), 0)

drone_2021 <- drone_2021 |> select(Block, Plot, mean_count)
#colnames(drone_2021) <- colnames(manual_dens_2021)


comp_21 <- merge(drone_2021,manual_dens_2021, by = c("Plot", "Block"), all.x = TRUE)
colnames(comp_21)



colnames(comp_21) <- c("Plot", "Block", "Drone", "Manual") 
comp_21$Manual <- as.numeric(comp_21$Manual)

summary(comp_21)

comp_21 <- comp_21 |> pivot_longer(cols = c("Drone", "Manual"), names_to = "Method", values_to = "Count")

ggplot(data = comp_21, aes(Plot, Count, fill = Method))+
   stat_summary(geom = "bar", fun = mean, position = "dodge", color = "black", width = 0.5) +
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2, position = position_dodge(0.5))+
  theme_classic() + 
    scale_y_continuous(expand = c(0, 0), limits = c(0, 11)) +  xlab( "Treatment") +
  ylab("Number of Mounds Detected")+
  theme(legend.position = c(0.8,0.75),
        text=element_text("serif", size=12), #change font size of all text
        axis.text=element_text("serif", size=12, colour = "black"), #change font size of axis text
        axis.title=element_text("serif", size=12), #change font size of axis titles
        plot.title=element_text("serif", size=12), #change font size of plot title
        legend.text=element_text("serif", size=12), #change font size of legend text
        legend.title=element_text("serif", size=12))+
  scale_fill_manual(values = c("grey28", "lightgrey"))+
  scale_x_discrete(labels = c("Burned:Excluded", "Burned:Present", "Unburned:Excluded", " Unburned:Present"))

aov21 <- aov(comp_21$Count~ comp_21$Method+ comp_21$Plot + comp_21$Method*comp_21$Plot)
summary(aov21)



```
Do the same as above but only for mounds measuring greater than 2m in diameter

```{r}

manual_large[manual_large=="BS"] <- "BR"
manual_large[manual_large=="US"] <- "UR"
comp_21_large <- merge(drone_2021, manual_large, by = c("Plot", "Block"), all.x = TRUE)
colnames(comp_21_large)


colnames(comp_21_large) <- c("Plot", "Block", "Drone", "Manual") 


comp_21_large <- comp_21_large |> pivot_longer(cols = c("Drone", "Manual"), names_to = "Method", values_to = "Count")

ggplot(data = comp_21_large, aes(Plot, Count, fill = Method))+
   stat_summary(geom = "bar", fun = mean, position = "dodge", color = "black", width = 0.5) +
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2, position = position_dodge(0.5))+
  theme_classic() + 
    scale_y_continuous(expand = c(0, 0), limits = c(0, 11)) +
    theme(title = element_text("serif"), text=element_text("serif", size=12), #change font size of all text
        axis.text=element_text("serif", size=12, colour = "black"), #change font size of axis text
        axis.title=element_text("serif", size=12), #change font size of axis titles
        plot.title=element_text("serif", size=12), #change font size of plot title
        legend.text=element_text("serif", size=12), #change font size of legend text
        legend.title=element_text("serif", size=12))+
  xlab( "Treatment") +
  ylab("Number of Mounds Detected")+ 
    scale_y_continuous(expand = c(0, 0), limits = c(0, 11)) +
  theme(legend.position = c(0.8,0.75), text = element_text(size=12, family = "serif"))+
  scale_fill_manual(values = c("grey28", "lightgrey"))+
  scale_x_discrete(labels = c("Burned:Excluded", "Burned:Present", "Unburned:Excluded", " Unburned:Present"))

aov21_large <-aov(comp_21_large$Count~ comp_21_large$Method+ comp_21_large$Plot + 
                    comp_21_large$Method*comp_21_large$Plot)
summary(aov21_large)

```


Compare my manual area to my drone area with all disks

```{r}


#### Organize manual areas #####
man_area_21 <- read.csv("Manual_area_2021.csv")
man_area_21 <- man_area_21[man_area_21$Plot != "C",] # Remove controls
man_area_21 <- man_area_21  %>% group_by(Block, Plot) %>%
  summarise(total.area = sum(Disk.Area),
            se = sd(Disk.Area)/ sqrt(n())) # Summarize to create total disk area per plot

man_area_21 <- man_area_21  %>% replace(is.na(.), 0) #remove NAs
man_area_21 <- man_area_21 %>%
  mutate_all(~replace(., . == "BS","BR")) # I used different naming conventions so I'm switching S to R
man_area_21 <- man_area_21 %>%
  mutate_all(~replace(., . == "US","UR"))
man_area_21

#### Organize 2021 drone areas ####
drone_2021 <- disks[disks$Year== 2021,]
drone_2021 <- drone_2021 %>% group_by(Block, Plot) %>%
  summarise(total.area = sum(Shape_Area),
            se = sd(Shape_Area)/ sqrt(n()))
drone_2021 <- complete_combinations[complete_combinations$Year==2021,] %>% 
  left_join(drone_2021, by = c("Block", "Plot")) 

drone_2021 <- drone_2021 %>% replace(is.na(.), 0)


#### Merge together #####

drone_man21_comp <- merge(drone_2021, man_area_21, by = c("Block", "Plot"), all.x = TRUE)
drone_man21_comp <- drone_man21_comp |> select(Block, Plot, total.area.x, total.area.y)

colnames(drone_man21_comp) <- c("Block", "Plot","Drone", "Manual")
drone_man21_comp$Manual <- as.numeric(drone_man21_comp$Manual)
drone_man21_comp

#### Plot comparison between manual and drone ####
drone_man21_comp |> 
  pivot_longer(cols = c("Drone", "Manual"), names_to = "Method", values_to = "Area") |> 
  ggplot(aes(Plot, Area, fill = Method))+
   stat_summary(geom = "bar", fun = mean, position = "dodge", color = "black", width = 0.5) +
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2, position = position_dodge(0.5))+
  theme_classic()+
    theme(title = element_text("serif"), text=element_text("serif", size=12), #change font size of all text
        axis.text=element_text("serif", size=12, colour = "black"), #change font size of axis text
        axis.title=element_text("serif", size=12), #change font size of axis titles
        plot.title=element_text("serif", size=12), #change font size of plot title
        legend.text=element_text("serif", size=12), #change font size of legend text
        legend.title=element_text("serif", size=12))+
  xlab( "Treatment") +
  ylab(expression(paste("Average Total Disk Area per Plot   ", (m^2))))+ 
    scale_y_continuous(expand = c(0, 0), limits = c(0, 100))+
  theme(legend.position = c(0.8,0.75), text = element_text(size=12, family = "serif"))+
  scale_fill_manual(values = c("grey28", "lightgrey"))+
  scale_x_discrete(labels = c("Burned:Excluded", "Burned:Present", "Unburned:Excluded", " Unburned:Present"))

ano.df2 <- drone_man21_comp |> 
  pivot_longer(cols = c("Drone", "Manual"), names_to = "Method", values_to = "Area")

aov21_b <- aov(data = ano.df2, Area ~ Plot + Method + Plot*Method)
summary(aov21_b)

drone_man21_comp |> 
  pivot_longer(cols = c("Drone", "Manual"), names_to = "Method", values_to = "Area") |>
  group_by(Plot, Method) |>
  summarise(mean_area = mean(Area))

```

Compare my manual area to my drone area with the 250cm diameter cutoff


```{r}


#### Organize manual areas #####
man_area_21 <- read.csv("Manual_area_2021.csv")
man_area_21 <- man_area_21[man_area_21$Plot != "C",] # Remove controls
man_area_21 <- man_area_21[man_area_21$Disk.Diameter..cm.>=250,]
man_area_21 <- man_area_21  %>% group_by(Block, Plot) %>%
  summarise(total.area = sum(Disk.Area),
            se = sd(Disk.Area)/ sqrt(n())) # Summarize to create total disk area per plot

man_area_21 <- man_area_21  %>% replace(is.na(.), 0) #remove NAs
man_area_21 <- man_area_21 %>%
  mutate_all(~replace(., . == "BS","BR")) # I used different naming conventions so I'm switching S to R
man_area_21 <- man_area_21 %>%
  mutate_all(~replace(., . == "US","UR"))
man_area_21

#### Organize 2021 drone areas ####
drone_2021

#### Merge together #####

drone_man21_comp <- merge( drone_2021, man_area_21, by = c("Block", "Plot"), all.x = TRUE)
drone_man21_comp <- drone_man21_comp |> select(Block, Plot, total.area.x, total.area.y)
colnames(drone_man21_comp) <- c("Block", "Plot", "Drone", "Manual")
drone_man21_comp$Manual <- as.numeric(drone_man21_comp$Manual)
drone_man21_comp

#### Plot comparison between manual and drone with 250cm cut off####
drone_man21_comp |> 
  pivot_longer(cols = c("Drone", "Manual"), names_to = "Method", values_to = "Area") |>
  ggplot(aes(Plot, Area, fill = Method))+
   stat_summary(geom = "bar", fun = mean, position = "dodge", color = "black", width = 0.5) +
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2, position = position_dodge(0.5))+
  theme_classic()+
    theme(title = element_text("serif"),
          text=element_text("serif", size=12), #change font size of all text
        axis.text=element_text("serif", size=12, colour = "black"), #change font size of axis text
        axis.title=element_text("serif", size=12), #change font size of axis titles
        plot.title=element_text("serif", size=12), #change font size of plot title
        legend.text=element_text("serif", size=12), #change font size of legend text
        legend.title=element_text("serif", size=12))+
  xlab( "Treatment") +
  ylab(expression(paste("Average Total Disk Area per Plot   ", (m^2))))+ 
    scale_y_continuous(expand = c(0, 0), limits = c(0, 100))+
  theme(legend.position = c(0.8,0.75), text = element_text(size=12, family = "serif"))+
  scale_fill_manual(values = c("grey28", "lightgrey"))+
  scale_x_discrete(labels = c("Burned:Excluded", "Burned:Present", "Unburned:Excluded", " Unburned:Present"))

drone_man21_comp |> 
  pivot_longer(cols = c("Drone", "Manual"), names_to = "Method", values_to = "Area") |>
  group_by(Plot, Method) |>
  summarise(mean_area = mean(Area))

ano.df1 <- drone_man21_comp |> 
  pivot_longer(cols = c("Drone", "Manual"), names_to = "Method", values_to = "Area") 
  
aov21_a <- aov(data = ano.df1, Area ~ Method + Plot + Method*Plot)
summary(aov21_a)

```


GLMM for density

```{r}

library(rstatix)
density

dens.mod <- glm(data = density, mean_count ~ Burned + Rodents + Year + Block, family = "poisson")

summary(dens.mod)
density$obs <- 1:length(density$mean_count)

```



