---
title: "Ant Mound eCognition Analysis"
author: "R Pienaar"
date: "`r Sys.Date()`"
output:
  word_document: default
  pdf_document: default
---


load packages
```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggpubr)
library(patchwork)
library(ggh4x)
library(Rmisc)
library(lme4)
library(stringr)

```



Input drone imagery counts from 2021
```{r}

d21 <- read.csv("2021_mound_per_plot.csv") ### Individual mounds from 2021 drone imagery

d21_d <- d21 %>%
  group_by(Burned, Rodents, Block) %>%
  dplyr::summarise(Count = n_distinct(OBJECTID)) #summarize the count by block and plot

all_combinations <- tidyr::expand(d21, Burned, Rodents, Block, ID = NULL)

d21_d <- dplyr::left_join(all_combinations, d21_d, by = c("Burned", "Rodents", "Block"))

d21_d[is.na(d21_d)]<- 0

d21_d %>%
  dplyr::group_by(Burned, Rodents) %>%
  dplyr::summarize(mean_count = mean(Count))

#### Plot 2021 ant mounds from drone imagery
ggplot(data = d21_d, aes(x = Burned, y = Count, fill = Rodents))+
  stat_summary(geom = "bar", fun = mean, position = "dodge", color = "black", width = 0.5) +
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2, position = position_dodge(0.5))+
  ylab("Disks per 30mx30m plot")+
  theme(axis.text.y = element_text(color = "black"))+
  theme(axis.text.x = element_text(color = "black"))+
  xlab("Fire Treatment")+
  theme_light()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_rect(fill="transparent"))+
  theme(plot.title = element_text(hjust = 0.5), text = element_text(size=15))+
  scale_fill_manual(values = c("darkseagreen","darkorchid"),labels=c('Rodents Excluded', 'Rodents Allowed'))+
  scale_x_discrete(labels= c("Burned", "Unburned"), )+
  theme(legend.position = "right", aspect.ratio = 1)




```

Input manual counts from 2021
```{r}

d21_m <- read.csv("Density_manual.csv")

#Plot manually counted data from 2021
ggplot(data = d21_m, aes(x = Burned, y = Count, fill = Rodents))+
  stat_summary(geom = "bar", fun = mean, position = "dodge", color = "black", width = 0.5) +
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2, position = position_dodge(0.5))+
  ylab("Disks per 30mx30m plot")+
  xlab(NULL)+
  theme(axis.text.y = element_text(color = "black"))+
  theme(axis.text.x = element_text(color = "black"))+
  theme_light()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_rect(fill="transparent"))+
  theme(plot.title = element_text(hjust = 0.5), text = element_text(size=15))+
  scale_fill_manual(values = c("darkseagreen","darkorchid"),labels=c('Rodents Excluded', 'Rodents Allowed'))+
  scale_x_discrete(labels= c("", ""), )+
  theme(legend.position = "none", aspect.ratio = 1)

```


Combine drone and manual counts for 2021
```{r}

d21_d
d21_m

d21_d$Block <- as.factor(d21_d$Block)
d21_m$Block <- as.factor(d21_m$Block)

colnames(d21_m)[colnames(d21_m) == "Count"]<- "Count_m"

d_both <- cbind(d21_d, d21_m$Count_m)

colnames(d_both)[colnames(d_both) == "d21_m$Count_m"]<- "Count_manual"
colnames(d_both)[colnames(d_both) == "Count"]<- "Count_drone"

d_both %>% unite("ID", Burned, Rodents, remove = FALSE) %>% 
  pivot_longer(cols = c(Count_drone, Count_manual), names_to = "type", values_to = "Count") %>%
ggplot(aes(x= ID, y = Count, fill = type))+
  #geom_bar(stat = "identity", position = position_dodge())+
  stat_summary(geom = "bar", fun = mean, position = "dodge", color = "black", width = 0.5) +
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2, position = position_dodge(0.5))+
  ylab("Disks per 30mx30m plot")+
  xlab("Treatment")+
  labs(fill = "Method")+
 theme_bw()+
  theme(panel.background = element_rect(fill="transparent"), panel.border = element_blank())+
  theme(plot.title = element_text(hjust = 0.5), text = element_text(size=12, family = 'serif'))+
  theme(legend.position = "right", aspect.ratio = 1)+
  theme(axis.line = element_line(color = 'black'))+ 
  scale_x_discrete(labels= c("BN", "BS", "UN", "US"), )+
  scale_fill_manual(values = c("darkseagreen","darkorchid"),labels=c('Drone count', 'Manual count'))+
  theme(legend.position = "right", aspect.ratio = 1)+ 
    scale_y_continuous(expand = c(0, 0), limits = c(0, 11)) 

```

A paired t-test showing the difference for counts between each treatment
```{r}

t.test(data = d_both, d_both$Count_drone, d_both$Count_manual, paired = TRUE)

d_both_long <- d_both |> pivot_longer(cols = c("Count_drone", "Count_manual"), names_to = "Method", values_to = "Count")
d_both_long

d_both_long$Method <- as.factor(d_both_long$Method)
d_both_long$Burned <- as.factor(d_both_long$Burned)
d_both_long$Rodents <- as.factor(d_both_long$Rodents)

levels(d_both_long$Block)

summary(aov(data = d_both_long, Count ~ Method)) #detection rate differences

man_drone_diff21 <- d_both
man_drone_diff21$diff <- man_drone_diff21$Count_manual-man_drone_diff21$Count_drone

summary(aov(data = man_drone_diff21, diff~Burned + Rodents))

```


Input drone imagery comparisons across all years
```{r}

drone_df <- read.csv("year_comp.csv")

head(drone_df)
colnames(drone_df) <- c("OID_", "Block", "Plot", "Burned", "Rodents", 
    "2011", "2016",  "2021", "2023")


drone_df <- drone_df %>% unite("ID", Burned, Rodents, remove = FALSE) %>% 
  pivot_longer(cols = c("2011", "2016", "2021", "2023"), names_to = "Year", 
               values_to = "Count")

all_combinations_wyear <- tidyr::expand(drone_df, Plot, Block, Year, ID = NULL)

#drone_df$Year <- as.numeric(drone_df$Year)

drone_df <- drone_df[drone_df$Count==1,] |>
  group_by(Plot, Year, Block) |> 
  dplyr::summarise(mean_count = mean(n())) 

drone_df <- left_join(all_combinations_wyear,drone_df, by = c("Plot", "Block", "Year"))
drone_df[is.na(drone_df)] <- 0

inc_df <- drone_df[drone_df$Year != 2023,]

inc_df |>
ggplot( aes(x=Plot, y = mean_count, fill = Year)) +
  #geom_col(position = position_dodge())+
  stat_summary(geom = "bar", fun = mean, position = "dodge", color = "black", width = 0.5) +
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2, position = position_dodge(0.5))+
  ylab("Disks per 30mx30m plot")+
  xlab("Fire Treatment")+
 theme_bw()+
  theme(panel.background = element_rect(fill="transparent"), panel.border = element_blank())+
  theme(plot.title = element_text(hjust = 0.5), text = element_text(size=12, family = 'serif'))+
  theme(legend.position = "right", aspect.ratio = 1)+
  theme(axis.line = element_line(color = 'black'))+
  #scale_x_discrete(labels= c("BN", "BS", "UN", "US"), )+
  theme(legend.position = "right", aspect.ratio = 1)+ 
    scale_y_continuous(expand = c(0, 0), limits = c(0, 6))+
  scale_fill_discrete(labels=c('2011', '2016', "2021", "2023"))

drone_df$Burned <- substr(drone_df$Plot,1,1)
drone_df$Rodents <- substr(drone_df$Plot,2,2)


  year_aov <- aov(mean_count~as.factor(Year)+Burned+ Rodents+ (1|Block), data = drone_df)
  summary(year_aov)
  TukeyHSD(year_aov)
  
```
This shows the difference between the total counts in each year, however the rodent effect is inflated because UN starts so much higher than the other treatments


Plot the density of the manually collected against the density of the drone collected
```{r}

d_both |> unite("ID", Burned, Rodents, sep = "", remove = FALSE) |>
  ggplot(aes(Count_drone, Count_manual, col = ID))+
  geom_point()+
  geom_smooth(method = "lm")+
  stat_regline_equation(aes(label = after_stat(rr.label)), label.x = 0,
                        label.y = c(15,14,13,12))+
    ylab("Manual count of ant mounds per 30mx30m plot")+
  xlab("Drone count of ant mounds per 30mx30m plot")+
  theme_bw()+
  theme(panel.background = element_rect(fill="transparent"), panel.border = element_blank())+
  theme(plot.title = element_text(hjust = 0.5), text = element_text(size=12, family = 'serif'))+
  theme(legend.position = "right", aspect.ratio = 1)+
  theme(axis.line = element_line(color = 'black'))

```

Plot a change in ant mounds over years
```{r, warning=FALSE}


inc_df$Year <- as.numeric(inc_df$Year)

inc_df |>
  group_by(Year, Plot) |>
  dplyr::summarise(ave = mean(mean_count),
                   se = sd(mean_count) / sqrt(n())) |>
    ggplot(aes(x=Year, y = ave, color = Plot))+
    geom_point() +
    geom_ribbon(aes(ymin = ave - se, ymax = ave + se, fill = Plot), alpha = 0.1) +  
    geom_smooth(aes(group = Plot), alpha = 0.1) +
    labs(title = "Line Plot with Standard Error",
     x = "Year",
     y = "Average",
     color = "Plot") +
    theme_bw()+
    theme(panel.background = element_rect(fill="transparent"), panel.border = element_blank())+
    theme(plot.title = element_text(hjust = 0.5), text = element_text(size=12, family = "serif"))+
    theme(legend.position = "right", aspect.ratio = 1)+
    theme(axis.line = element_line(color = 'black'))





```

Plot Net Change from 2011 to 2023, this shows the increase in number of mounds so that the starting number doesn't affect the results
```{r}

reshaped <- drone_df
reshaped <- reshaped %>%
  pivot_wider(names_from = Year, values_from = mean_count)

# Calculate the difference between the values for 2021 and 2011
reshaped <- reshaped %>%
  mutate(increase = `2021` - `2011`) 

reshaped <- reshaped %>%
  mutate(decrease = `2023` - `2021`) %>%
  select(Plot, increase, decrease)
reshaped$decrease <- reshaped$decrease*-1
  
#Increase from 2011 to 2021
reshaped |>
  ggplot( aes(x = increase, y = Plot)) +
  stat_summary(geom = "bar", fun = mean, color = "black", fill = "skyblue", width = 0.5) +
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2)+
  #geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs( x = "Increase in mounds from 2011 to 2021",
       y = "Plot") +
  #geom_text(aes(label = scales::number(difference, accuracy = 0.1)), hjust = -0.2) +
  theme_bw()+
  theme(panel.background = element_rect(fill="transparent"), panel.border = element_blank())+
  theme(plot.title = element_text(hjust = 0.5), text = element_text(size=12, family = 'serif'))+
  theme(legend.position = "right", aspect.ratio = 1)+
  theme(axis.line = element_line(color = 'black'))+
  scale_x_continuous(expand = c(0, 0), limits = c(0, 5))

#Decrease from 21 to 23
reshaped |>
  ggplot( aes(x = decrease, y = Plot)) +
  stat_summary(geom = "bar", fun = mean, color = "black", fill = "pink2", width = 0.5) +
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2)+
  #geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs( x = "Decrease from 2021 to 2023",
       y = "Plot") +
  #geom_text(aes(label = scales::number(difference, accuracy = 0.1)), hjust = -0.2) +
  theme_bw()+
  theme(panel.background = element_rect(fill="transparent"), panel.border = element_blank())+
  theme(plot.title = element_text(hjust = 0.5), text = element_text(size=12, family = 'serif'))+
  theme(legend.position = "right", aspect.ratio = 1)+
  theme(axis.line = element_line(color = 'black'))


#Actual stats instead of just visualizations
# Calculate the difference between the values for 2023 and 2011

reshaped$Burned <- substr(reshaped$Plot, 1, 1)
reshaped$Rodents <- substr(reshaped$Plot, 2, 2)

inc_aov <- aov(data= reshaped, increase~ Burned + Rodents +Burned*Rodents)
summary(inc_aov)
TukeyHSD(inc_aov)

dec_aov <- aov(data= reshaped, decrease ~ Burned + Rodents +Burned*Rodents)
summary(dec_aov)
TukeyHSD(dec_aov)

```



Joshs Data from 2016
```{r}

manual_d_16 <- read.csv("C:/Users/ryanp/Documents/Dissertation/Chapter1_Ants/Josh data/Ant mound density.csv")
head(manual_d_16)
colnames(manual_d_16)[1]<- "Block"
colnames(manual_d_16)[3]<- "Burned"
colnames(manual_d_16)[4]<- "Rodents"
manual_d_16$Burned <- toupper(manual_d_16$Burned)
manual_d_16$Rodents <- toupper(manual_d_16$Rodents)

```

Compare 2016 and 2021 manuals to their classifications

```{r}

d_both <- merge(d_both, manual_d_16, by = c("Block", "Burned", "Rodents"), all.x = TRUE)
d_both <- d_both[, c("Block", "Burned", "Rodents", "Count_drone", "Count_manual", "density")]
colnames(d_both)[colnames(d_both) == "density"]<- "Count_manual_16"

d_both %>% unite("ID", Burned, Rodents, remove = FALSE) %>% 
  pivot_longer(cols = c(Count_manual, Count_manual_16), names_to = "Year", values_to = "n") %>%
ggplot(aes(x= ID, y = n, fill = Year))+
  #geom_bar(stat = "identity", position = position_dodge())+
  stat_summary(geom = "bar", fun = mean, position = "dodge", color = "black", width = 0.5) +
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2, position = position_dodge(0.5))+
  ylab("Manually counted isks per 30mx30m plot")+
  xlab("Fire Treatment")+
 theme_bw()+
  theme(panel.background = element_rect(fill="transparent"), panel.border = element_blank())+
  theme(plot.title = element_text(hjust = 0.5), text = element_text(size=12, family = 'serif'))+
  theme(legend.position = "right", aspect.ratio = 1)+
  theme(axis.line = element_line(color = 'black'))+ 
  scale_x_discrete(labels= c("BN", "BS", "UN", "US"), )+
  scale_fill_manual(values = c("darkseagreen","darkorchid"),labels=c('2021', "2016"))+
  theme(legend.position = "right", aspect.ratio = 1)+ 
    scale_y_continuous(expand = c(0, 0), limits = c(0, 11)) 

```

Analysis for differences between 2016 and 2021 manual counts

```{r}

t.test(d_both$Count_manual, d_both$Count_manual_16, paired = TRUE)


d_both$man_diff <- d_both$Count_manual-d_both$Count_manual_16
d_both$Plot <- str_c(d_both$Burned,d_both$Rodents, sep = "")

d_both |> select(Plot, man_diff) |>
  ggplot( aes(x = man_diff, y = Plot)) +
  stat_summary(geom = "bar", fun = mean, position = "dodge", color = "black", fill = "skyblue", width = 0.5) +
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2, position = position_dodge(0.5))+
  #geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs( x = "Difference in mounds between 2011 and 2023",
       y = "Plot") +
  #geom_text(aes(label = scales::number(difference, accuracy = 0.1)), hjust = -0.2) +
  theme_bw()+
  theme(panel.background = element_rect(fill="transparent"), panel.border = element_blank())+
  theme(plot.title = element_text(hjust = 0.5), text = element_text(size=12, family = 'serif'))+
  theme(legend.position = "right", aspect.ratio = 1)+
  theme(axis.line = element_line(color = 'black'))+
  scale_x_continuous(expand = c(0, 0), limits = c(-5, 5))

### This needs to be compared to the drone 2021 vs 2016

reshaped |>
  ggplot( aes(x = increase, y = Plot)) +
  stat_summary(geom = "bar", fun = mean, position = "dodge", color = "black", fill = "lightgreen", width = 0.5) +
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2, position = position_dodge(0.5))+
  #geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs( x = "Difference in mounds between 2021 and 2016",
       y = "Plot") +
  #geom_text(aes(label = scales::number(difference, accuracy = 0.1)), hjust = -0.2) +
  theme_bw()+
  theme(panel.background = element_rect(fill="transparent"), panel.border = element_blank())+
  theme(plot.title = element_text(hjust = 0.5), text = element_text(size=12, family = 'serif'))+
  theme(legend.position = "right", aspect.ratio = 1)+
  theme(axis.line = element_line(color = 'black'))+
  scale_x_continuous(expand = c(0, 0), limits = c(-5, 5))

testa <- d_both |> select(Plot, man_diff)

testb <- drone_df |> 
  pivot_wider(names_from = Year, values_from = mean_count) |>
  mutate(difference = `2021`-`2016`) |>
  select(Plot, difference)
  
testb[19,1] <-"UN"
testb[20,1] <-"US"

testb[is.na(testb)]<- 0
testb <- as.data.frame(testb)

testa <- testa[order(testa$Plot),]
testb <- testb[order(testb$Plot),]

testc <- testa
testc$newdf <- testb$difference

t.test(testc$man_diff,testc$newdf)

testc |> pivot_longer(cols = c(man_diff, newdf), names_to = "Method", values_to = "Count") |> 
ggplot( aes(x = Plot, y = Count, fill = factor(Method, levels = c('newdf', "man_diff")))) +
  stat_summary(geom = "bar", fun = mean, position = "dodge", color = "black", width = 0.5) +
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2, position = position_dodge(0.5))+
  ylab("Change in mounds")+
  xlab("Treatment")+
 theme_bw()+
  theme(panel.background = element_rect(fill="transparent"), panel.border = element_blank())+
  theme(plot.title = element_text(hjust = 0.5), text = element_text(size=12, family = 'serif'))+
  theme(legend.position = "right", aspect.ratio = 1)+
  theme(axis.line = element_line(color = 'black'))+ 
  scale_x_discrete(labels= c("BN", "BS", "UN", "US"), )+
  scale_fill_manual("Method", values =  c("darkseagreen","darkorchid"),labels=c('Drone count', "Manual count"))+
  theme(legend.position = "right", aspect.ratio = 1)+ 
    scale_y_continuous(expand = c(0, 0), limits = c(0, 9)) 

```

So there are differences in the change in ant mounds detected between 2016 and 2021 by drone and manual counting, but what if I can demonstrate that they show the same trend?

```{r}

testc$Burned <- substr(testc$Plot,1,1)
testc$Rodents <- substr(testc$Plot,2,2)

testc_long<- testc |> pivot_longer(cols = c("man_diff", "newdf"), 
                                   names_to = "Method", 
                                   values_to = "Count") 
 summary(aov(data = testc_long, Count~ Method))


man_aov <- aov(data = testc, man_diff~Burned+Rodents)
dr_aov <- aov(data = testc, newdf~Burned+Rodents)

summary(man_aov)
summary(dr_aov)

```

Connect github
```{r}

install.packages("usethis")
library(usethis)
use_github()
use_git()

usethis::use_git()

usethis::use_github()	

```



